---
title: "model_comparisons"
author: "Andrea Stocco"
date: "5/24/2021"
output:
  html_document:
    code_folding: hide
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(magrittr)
library(ggplot2)
library(scales)
library(ggthemes)
library(ggsci)
library(kableExtra)
library(xtable)
library(matlab)
library(R.matlab)
library(MCMCpack)
library(ggrepel)
library(patchwork)    
```
## Resting state 

First, some constants: Models and their associated colors

```{r}
model.names <- c("Common Model", 
                 "Hierarchical 1", 
                 "Hierarchical 2",
                 "Hierarchical 3",
                 "Hub-and-spoke PFC", 
                 "Hub-and-spoke BG",
                 "Hub-and-spoke Temporal"
  )

model_colors <- c("red3", 
                  "palegreen4", "palegreen3", "palegreen2",
                  #"goldenrod4", "goldenrod3", "goldenrod2",
                  "deepskyblue4", "deepskyblue3", "deepskyblue2")

```

And now, two utility functions, one to load the data of a BMS data structure, and one to calculate Dirichlet distributions fro the BMS's $\alpha$ values.

```{r}
load.data <- function(file, 
                      models=model.names, 
                      task="Resting State") {
  res <- NULL
  bms <- readMat(file)
  data <- bms$BMS
  # Likelihood
  
  like <- data[[1]][[1]][[4]]
  # RFX analysis
  
  alpha <- data[[1]][[1]][[5]][,,1]$alpha
  expect <- data[[1]][[1]][[5]][,,1]$exp.r
  exceed <- data[[1]][[1]][[5]][,,1]$xp
  pexceed <- data[[1]][[1]][[5]][,,1]$pxp
  partial <- data.frame(Model = models, Alpha = c(alpha),
                        Likelihood = c(like), Expected = c(expect),
                        Exceedance = c(exceed), PExceedance = c(pexceed),
                        Task = task)
  
  if (is.null(res)) {
    res <- partial
  } else {
    res <- merge(res, partial, all=T)
  }
  res
}

generate.exp <- function(alphas, 
                         mnames = paste("M", 1:length(alphas), sep=""), 
                         res = 0.001) {
  p <- seq(0, 1, res)
  ii <- 1:length(alphas)
  dd <- NULL
  
  for (i in ii) {
    alpha <- alphas[i]
    others <- sum(alphas[ii != i])  # Dirichlet dist is agglomerative
    #print(c(alpha, others))   # Debug
    
    # Create the appropriate instantiation of the Dirichlet
    # Functions with two alpha values (each model vs. the rest)
    f <- function(x) {ddirichlet(c(x,  1 -x), alpha=c(alpha, others))}
    vf <- Vectorize(f)
    d <- vf(p)
    d[is.na(d)] <- 0
    dd <-cbind(dd, d)
    #print(dim(dd))
  }
  colnames(dd) <- mnames
  
  # Return
  dd
}
```

Now, let's load the data:

```{r}
comparisons <- load.data("comparisons/BMS.mat")
```

We create the Relatie Log-Likelihood measure and print the full table of measures:

```{r}
minL <- min(comparisons$Likelihood)

comparisons <- comparisons %>%
  mutate(RelativeLikelihood = Likelihood - minL)

comparisons %>%
  kable() %>%
  kable_styling(bootstrap_options = c("hover", "striped"))
```

## Probability density distributions

```{r}
alphas <- comparisons$Alpha
dists <- as.data.frame(generate.exp(alphas)/1000)


names(dists) <- model.names

wdists <- as_tibble(dists) %>%
  add_column(Task = "Resting State",
             Probability = seq(0, 1, 1/1000))

ldists <- wdists %>%
  pivot_longer(cols = model.names,
                names_to = "Model",
               values_to = "Density") 
```

And now, plot:

```{r}
ldists$Density[ldists$Density == Inf] <- 0
ldist_summary <- ldists %>%
  group_by(Model) %>%
#  mutate(dXdY = Probability * Density * 1000) %>%
  summarise(#Expected = mean(dXdY),
            MaxDensity = max(Density))

comparisons <- inner_join(comparisons, ldist_summary, keep=F)

densityp <- ggplot(ldists, aes(x=Probability, 
                               y = Density, 
                               fill=Model, 
                               col=Model)) +
  geom_polygon(alpha=.5) +
  scale_color_manual(values = model_colors) +
  scale_fill_manual(values = model_colors) +
  geom_text_repel(data=comparisons, aes(x=Expected, 
                                        y=MaxDensity, 
                                        label=percent(Exceedance, accuracy = 0.1)),
                  col="black") +
  ggtitle("Resting State") +
  theme_pander()

densityp 
```


## Exceedance probabilities

And now, the exceedance probabilities:

```{r}
ggplot(comparisons, aes(x="", y = Exceedance, fill=Model, col=Model)) +
  geom_bar(stat = "identity", col="grey", width=1, alpha=.8) +
  coord_polar("y", start=0) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  scale_color_manual(values = model_colors) +
  scale_fill_manual(values = model_colors) +
  ggtitle("Resting State") +
  xlab("Exceedance Probabilities") +
  ylab("Resting State") +
  geom_text_repel(aes(label=percent(Exceedance, .1)), 
                  col="white",
                  position=position_stack(vjust=0.5), 
                  direction="both",
                  size=3) +
  theme_pander() +
  theme(axis.ticks = element_blank(), 
        axis.text.y = element_blank(),
        axis.text.x = element_blank()) 

```

Same plot, but, this time, as stacked horizontal bars (like in the NeuroImage paper)

```{r fig.width=6, fig.height=2}
exceedancep <- 
  ggplot(comparisons, aes(x=Task, y = Exceedance, fill=Model, col=Model)) +
  geom_bar(stat = "identity",  
           width = .5, alpha = .5) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  scale_color_manual(values = model_colors) +
  scale_fill_manual(values = model_colors) +
  ggtitle("Resting-State fMRI") +
  xlab("Task") +
  ylab("Exceedance Probabiliy") +
  coord_flip() +
  geom_text_repel(aes(label=percent(Exceedance, .1)), 
                  col="black",
                  position=position_stack(vjust=0.5), 
                  direction="both",
                  size=3) +
  theme_pander() +
  theme(legend.position = "bottom")

exceedancep
```

And here we have a paper-worth combined figure:

```{r, fig.width=5, fig.height=4}
densityp + ggtitle("")  + 
  theme(legend.position = "bottom") + 
  guides(fill=guide_legend(ncol=2),
         col=guide_legend(ncol=2))
```

```{r fig.width=5, fig.height=4}

logLp <- ggplot(comparisons, aes(x=Model, 
                                 y = RelativeLikelihood, 
                                 fill=Model,
                                 col=Model)) +
  geom_col(
           width = .9, alpha = .5) +
  scale_color_manual(values = model_colors) +
  scale_fill_manual(values = model_colors) +
  ggtitle("Resting-State fMRI") +
  xlab("Task") +
  ylab("Relative Log-Likelihood") +
  #coord_flip() +
  geom_text(aes(label=round(RelativeLikelihood,1),
                y = RelativeLikelihood + 150), 
            col="black",
#            position = position_stack(vjust=c(1)),
            direction="both",
            size=3) +
  theme_pander() +
  theme(axis.text.x = element_text(angle=45, hjust = c(1)),
        legend.position = "NA")

logLp
```
