---
title: "Resting State vs. Task Based Parameters"
author: "Andrea Stocco"
date: "5/24/2021"
output:
  html_document:
    code_folding: hide
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(magrittr)
library(ggplot2)
library(scales)
library(ggthemes)
library(ggsci)
library(kableExtra)
library(xtable)
library(matlab)
library(R.matlab)
library(MCMCpack)
library(ggrepel)
```
# Resting state vs. task-based parameters

First, let's define some basic variables.

```{r}
folders <- c("Emotion", "Gambling", "Language", 
             "Relational", "Social", "WM")

task_names <- c("Emotion Recognition", 
                "Incentive Processing", "Language and Math", 
                "Relational Reasoning", "Social Cognition", 
                "Working Memory")

regions <- c("Action", "LTM", "Perception", "Procedural", "WM")
```

## Data loading and preparation

The data is contained in a series of task- and paradigm-specific matrices. 

```{r}
res <- NULL
for (task in folders) {
  vals <- scan(file = paste("../tfMRI/all_architectures/", 
                            task, 
                            paste(task, "_avg_A.txt", sep=""),
                            sep="/"),
               sep=",")
  if (is.null(res)) {
    res <- vals
  } else {
    res <- rbind(res, vals)
  }
}

vals <- scan("Rest1_avg_A.csv", sep=",")
res <- rbind(res, vals)

```

Now, let's remove the missing connections and turn the results into a tibble

```{r}
connection_names <- c(outer(regions, regions, FUN = function(x, y) {paste(x, "->", y)}))
colnames(res) <- connection_names
nres <- res[,colMeans(res) != 0]
rownames(nres) <- c(task_names, "Resting state")
nres <- as_tibble(nres)
```

And, finally, let's put the tibble in long form, so it can be visualized with `ggplot`.

```{r}
wres <- nres %>%
  add_column(Task = c(task_names, "Resting state") )

wres <- wres %>%
  mutate(TaskVsRest = if_else(Task == "Resting state", "Rest", "Task"))

lres <- wres %>%
  pivot_longer(colnames(nres), names_to = "Connection", values_to = "Value")
```

Now, we wnat to change the order of factors in "Task" so that the Resting State is always listed last.

```{r}
lres$Task <- factor(lres$Task,
                    levels = c(task_names, "Resting state"))
```

## Results

Here is the visuaklization

```{r fig.width=6, fig.height=6}

ggplot(lres, aes(y=Connection, x=Value, col=Task)) +
  geom_point(alpha=0.6, position=position_jitter(height=0.2), size=5) +
  stat_summary(data = filter(lres, TaskVsRest=="Task"), 
               fun.data = mean_sdl, 
               geom="errorbar", 
               col="navy", 
               width=0.25) +
  scale_color_manual(values = c("deepskyblue1", "dodgerblue", "dodgerblue1", "dodgerblue2", "dodgerblue3", "dodgerblue4", "red")) +
  scale_shape_manual(values=c(21, 22)) +
  ylab("Connection") +
  xlab("Effective Connectivity (Hz)") +
  ggtitle("Connectivity Values\nAcross Tasks and Rest Data") +
  labs(col="Task") +
  theme_pander() +
  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position="right")

ggsave("figure11.png")
```

## Are there significant differences?

We can use a simple ANOVA to see whether there is a significant effect of task:

```{r}
summary(aov(Value ~ (Connection * TaskVsRest), lres)) %>%
  xtable() %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## How similar is Resting-State to Task based data?

```{r}
mres <- as.matrix(nres)
R <- cor(t(mres))
R[R==1] <- NA
tR <- as_tibble(R)
names(tR) <- c(task_names, "Resting state")
wR <- tR %>%
  add_column(From = c(task_names, "Resting state"))
lR <- wR %>%
  pivot_longer(!From, names_to="To", values_to = "Correlation")

wR   %>%
  xtable() %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

lR   %>%
  xtable() %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

lR$From <- factor(lR$From, levels=c(task_names, "Resting state"))
lR$To <- factor(lR$To, levels=c(task_names, "Resting state"))

```


And now, visualize:


```{r}
ggplot(lR, aes(x=From, y=To, fill=Correlation)) +
  geom_tile(col="white") +
 # scale_fill_viridis_b(option="plasma") +
  theme_pander() +
  scale_fill_viridis_c(option = "plasma") +
  ggtitle("Parameter Correlations Across Tasks and Rest") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

And now, what is the mean correlation between one task and all the others? 

```{r}

r2z <- function(r) {
  ((1 + r)/(1-r))/2
}

z2r <- function(z) {
  (2 * z - 1) / (2 * z + 1)
}

lR <- lR %>%
  mutate(Z = r2z( Correlation ))

withinTaskMeans <- lR %>%
  filter(To != "Resting state") %>%
  filter(From != "Resting state") %>%
  group_by(To) %>%
  summarise(meanZ=mean(Z, na.rm=T)) %>%
  mutate(meanR = z2r(meanZ))

withinTaskMeans %>%
  xtable() %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


restingStateMeans <- lR %>%
  filter(From == "Resting state") %>%
  filter(To != "Resting state") %>%
  #group_by(To) %>%
  summarise(meanZ= mean(Z, na.rm=T)) %>%
  mutate(meanR = z2r(meanZ))

restingStateMeans %>%
  xtable() %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

The mean correlation between tasks parameters is `r z2r(mean(withinTaskMeans$meanZ))`, with range `r range(withinTaskMeans$meanR)`. The mean correlation of Resting State with task data is `r restingStateMeans$meanR`, which is well within the range.

## Multidimensional scaling

```{r}
fit <- cmdscale(dist(nres))
x <- fit[,1]
y <- fit[,2]
#plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2",
#  main="Metric MDS", type="n")
#text(x, y, labels = c(task_names, "Resting state"), cex=.7)

mds <- tibble(data.frame(X=x, Y=y, Paradigm = c(task_names, "Resting state")))
```

Let's visualize:

```{r}
ggplot(mds, aes(x=X, y=Y, col=Paradigm)) +
  geom_point(size=3) +
  xlim(-1, 2) +
  ylim(-1, 2) +
  xlab("Dimension 1") +
  ylab("Dimension 2") +
  scale_color_d3() +
  ggtitle("Multi-Dimensional Scaling\nof Connectivity Parameters") +
  geom_text_repel(aes(label = Paradigm)) +
  theme_pander() +
  coord_equal() +
  theme(legend.position = "NA")
ggsave("figure12.png")
```