---
title: "model_comparisons"
author: "Andrea Stocco"
date: "5/24/2021"
output:
  html_document:
    code_folding: hide
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(magrittr)
library(ggplot2)
library(scales)
library(ggthemes)
library(ggsci)
library(kableExtra)
library(xtable)
library(matlab)
library(R.matlab)
library(MCMCpack)
library(ggrepel)
```
## Resting state 

First, some constants: Models and their associated colors

```{r}
model.names <- c("Common Model", 
                 "Hierarchical 1", 
                 "Hierarchical 2",
                 "Hierarchical 3",
                 "Hub-and-spoke PFC", 
                 "Hub-and-spoke BG",
                 "Hub-and-spoke Temporal"
  )

model_colors <- c("red3", 
                  #"palegreen4", "palegreen3", "palegreen2",
                  "goldenrod4", "goldenrod3", "goldenrod2",
                  "deepskyblue4", "deepskyblue3", "deepskyblue2")

folders <- c("All", "Emotion", "Gambling", "Language", 
             "Relational", "Social", "WM")

task_names <- c("All Tasks Combined ", "Emotion Recognitin", 
                "Incentive Processing", "Language and Math", 
                "Relational Reasoning", "Social Cognition", 
                "Working Memory")
```

And now, two utility functions, one to load the data of a BMS data structure, and one to calculate Dirichlet distributions fro the BMS's $\alpha$ values.

```{r}
load.data <- function(file, 
                      models=model.names, 
                      task="Resting State") {
  bms <- readMat(file)
  data <- bms$BMS
  # Likelihood
  
  like <- data[[1]][[1]][[4]]
  # RFX analysis
  
  alpha <- data[[1]][[1]][[5]][,,1]$alpha
  expect <- data[[1]][[1]][[5]][,,1]$exp.r
  exceed <- data[[1]][[1]][[5]][,,1]$xp
  pexceed <- data[[1]][[1]][[5]][,,1]$pxp
  partial <- data.frame(Model = models, Alpha = c(alpha),
                        Likelihood = c(like), Expected = c(expect),
                        Exceedance = c(exceed), PExceedance = c(pexceed),
                        Task = task)
  
  partial
}

generate.exp <- function(alphas, 
                         mnames = paste("M", 1:length(alphas), sep=""), 
                         res = 0.001) {
  p <- seq(0, 1, res)
  ii <- 1:length(alphas)
  dd <- NULL
  
  for (i in ii) {
    alpha <- alphas[i]
    others <- sum(alphas[ii != i])  # Dirichlet dist is agglomerative
    #print(c(alpha, others))   # Debug
    
    # Create the appropriate instantiation of the Dirichlet
    # Functions with two alpha values (each model vs. the rest)
    f <- function(x) {ddirichlet(c(x,  1 -x), alpha=c(alpha, others))}
    vf <- Vectorize(f)
    d <- vf(p)
    d[is.na(d)] <- 0
    dd <-cbind(dd, d)
    #print(dim(dd))
  }
  colnames(dd) <- mnames
  
  # Return
  dd
}
```

Now, let's load the data:

```{r}
res <- NULL
  
for (task in folders) {
  comparisons <- load.data(paste(task, "BMS.mat", sep="/"),
                           task=task)
  if (is.null(res)) {
    res <- comparisons
  } else {
    res <- merge(res, comparisons, all=T)
  }
}
             
res %>%
  kable() %>%
  kable_styling(bootstrap_options = c("hover", "striped"))
```

## Probability density distributions

```{r}
all_dists <- NULL
for (task in folders) { 
  comparisons <- filter(res, Task == task)
  alphas <- comparisons$Alpha
  dists <- as.data.frame(generate.exp(alphas)/1000)
  
  names(dists) <- model.names
  
  wdists <- as_tibble(dists) %>%
    add_column(Task = task,
               Probability = seq(0, 1, 1/1000))
  
  ldists <- wdists %>%
    pivot_longer(cols = model.names,
                 names_to = "Model",
                 values_to = "Density") 
  
  if (is.null(all_dists)) {
    all_dists <- ldists
  }  else {
    all_dists <- rbind(all_dists, ldists)
  }
}
```

And now, plot:

```{r fig.width=7, fig.height=6}
all_dists$Density[all_dists$Density == Inf] <- 0
ggplot(filter(all_dists, Task!="All"), 
       aes(x=Probability, y = Density, 
           fill=Model, col=Model)) +
  geom_polygon(alpha = .5) +
  facet_wrap(~ Task, ncol=2) +
  coord_cartesian(ylim = c(0, 0.03)) +
  scale_color_manual(values = model_colors) +
  scale_fill_manual(values = model_colors) +
  ggtitle("Task-Based fMRI") +
  theme_pander() +
  theme(legend.position = "right")
```


## Exceedance probabilities

And now, the exceedance probabilities:

```{r}
ggplot(res, aes(x="", y = Exceedance, fill=Model, col=Model)) +
  geom_bar(stat = "identity", col="grey", width=1, alpha=.8) +
  coord_polar("y", start=0) +
  facet_wrap(~ Task) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2L)) +
  scale_color_manual(values = model_colors) +
  scale_fill_manual(values = model_colors) +
  ggtitle("Resting State") +
  xlab("Exceedance Probabilities") +
  ylab("Resting State") +
  geom_text_repel(aes(label=percent(Exceedance, .1)), 
                  col="white",
                  position=position_stack(vjust=0.5), 
                  direction="both",
                  size=3) +
  theme_pander() +
  theme(axis.ticks = element_blank(), 
        axis.text.y = element_blank(),
        axis.text.x = element_blank()) 

```

Same plot, but, this time, as stacked horizontal bars (like in the NeuroImage paper)

```{r}
ggplot(res, aes(x=Task, y = Exceedance, fill=Model, col=Model)) +
  geom_bar(stat = "identity", col="white", 
           width = .75, alpha = .5) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  scale_color_manual(values = model_colors) +
  scale_fill_manual(values = model_colors) +
  ggtitle("Task-based fMRI") +
  xlab("Task") +
  ylab("Exceedance Probabiliy") +
  coord_flip() +
  geom_text_repel(aes(label=percent(Exceedance, .1)), 
                  col="white",
                  position=position_stack(), 
                  direction="both",
                  size=3) +
  theme_pander()

```

# Example of fit

```{r, fig.height=7, fig.width=4}
pred <- read_delim("predicted.csv", delim=",",
                   col_names=c("Action", "LTM", "Perception", "Procedural", "WM"),
                   col_types = cols(
                     Action = col_double(),
                     LTM = col_double(),
                     Perception = col_double(),
                     Procedural = col_double(),
                     WM = col_double()
                   ))

# obs <- read_delim("responses.csv", delim=",",
#                   col_names=c("Action", "LTM", "Perception", "Procedural", "WM"),
#                   col_types = cols(
#                      Action = col_double(),
#                      LTM = col_double(),
#                      Perception = col_double(),
#                      Procedural = col_double(),
#                      WM = col_double()
#                    ))

obs <- read_tsv("responses.txt", 
                  col_names=c("Action", "LTM", "Perception", "Procedural", "WM"),
                  col_types = cols(
                     Action = col_double(),
                     LTM = col_double(),
                     Perception = col_double(),
                     Procedural = col_double(),
                     WM = col_double()
                   ))

time <- seq(1, nrow(pred)) * 0.72

lpred <- pred %>% 
  add_column(Time = time) %>%
  pivot_longer(cols=c("Action", "LTM", "Perception", "Procedural", "WM"),
                       names_to = "Region",
                       values_to = "BOLD")  %>%
  add_column(Type = "Predicted")

lobs <- obs %>% 
  add_column(Time = time) %>%
  pivot_longer(cols=c("Action", "LTM", "Perception", "Procedural", "WM"),
                       names_to = "Region",
                       values_to = "BOLD") %>%
  add_column(Type = "Observed")


lfit <- rbind(lobs, lpred)


ggplot(lfit, aes(x=Time, y=BOLD, col=Region, linetype=Type, alpha=Type)) +
  geom_line() +
  scale_alpha_manual(values=(c(0.35, 1))) +
  scale_linetype_manual(values=c("solid", "solid")) +
  scale_color_manual(values=c("purple", "orange3", "red", "chartreuse4", "darkturquoise")) +
  facet_wrap(~Region, ncol=1, scales = "free_y") +
  ylab("% BOLD Change") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  guides(color=F)
              
```

Just the timeseries:

```{r}
ggplot(filter(lfit, Type=="Observed"), 
       aes(x=Time, y=BOLD, col=Region, linetype=Type)) +
  geom_line(alpha=0.75) +
  #scale_alpha_manual(values=(c(0.35, 1))) +
  scale_linetype_manual(values=c("solid", "solid")) +
  scale_color_manual(values=c("purple", "orange1", "red", "chartreuse4", "darkturquoise")) +
  facet_wrap(~Region, ncol=3, scales = "free_y") +
  ylab("% BOLD Change") +
  theme_pander() +
  theme(legend.position = "NA") +
  guides(color=F)
```
